
---
title: "STAT 184 Final Project"
author: "Bernice Chen & Claudia Jung"
date: "December 17, 2019"
output: html_notebook
---


# Set Up

### Clean up environment and load packages
```{r}
# clean up the RStudio environment 
rm(list = ls())

# load all packages here: `mosaic`, `tidyverse`, `lubridate`, and all others used
library(mosaic)
library(tidyverse)
library(lubridate)
library(DataComputing)
library(ggplot2)
library(party)
```

### Data Intake

```{r}
### Load the data set into our RStudio environment
Starbucks <- read.csv("StarbucksinCaliforniaCityStats.csv")

#commands used to inspect the data set
names(Starbucks)
View(Starbucks)

#adding the State column
Starbucks <-
  Starbucks %>%
  mutate(State="California")
```

### Research Question

#### Is there a relationship between number of Starbucks and demographic characteristics of a city?

Factors that we are interested:

- Age

- Median Income

- Percentage of White population

- Location

### Analysis- Age

```{r}
#Some of the variables provided wasn't numerical variable, needed to be converted
Starbucks$Median.Age <- as.numeric(as.character(Starbucks$Median.Age))
View(Starbucks)
#Checking if it's successfully converted
is.numeric(Starbucks$Median.Age)
#Plotting number of Starbucks vs Age
Starbucks %>%
  ggplot(aes(x=Median.Age, y=Number.of.Starbucks)) + geom_point() + ylim(0,50)
```
### Analysis- Median Income
```{r}
#Some of the variables provided wasn't numerical variable, needed to be converted
#Getting rid of $ and ,
Starbucks <-
  Starbucks %>%
  mutate(Median.Household.Income=gsub(pattern = "[$,]", replacement = "", Median.Household.Income))
#converting it into numerical variable
Starbucks$Median.Household.Income <- as.numeric(as.character(Starbucks$Median.Household.Income))
View(Starbucks)
#Checking if it's successfully converted
is.numeric(Starbucks$Median.Household.Income)
Starbucks %>%
  ggplot(aes(x=Median.Household.Income, y=Number.of.Starbucks)) + geom_point() + ylim(0,50)


```

### Analysis- White Population
```{r}
#Some of the variables provided wasn't numerical variable, needed to be converted
Starbucks$Percentage.of.white.population <- as.numeric(as.character(Starbucks$Percentage.of.white.population))
View(Starbucks)
#Checking if it's successfully converted
is.numeric(Starbucks$Percentage.of.white.population)
#Plotting number of Starbucks vs Percentage of White Population
Starbucks %>%
  ggplot(aes(x=Percentage.of.white.population, y=Number.of.Starbucks)) + geom_point() +ylim(0,50)
```

### Top 10 and Population
```{r}
Starbucks %>%
  arrange(-Number.of.Starbucks) %>%
  select(City, Number.of.Starbucks, Median.Age, Median.Household.Income, Percentage.of.white.population) %>%
  head(10)

data("ZipGeography")
View(ZipGeography)

GroupByCity <-
  ZipGeography %>%
  group_by(CityName) %>%
  mutate(totalPopulation=sum(Population, na.rm=TRUE)) %>%
  distinct(CityName, .keep_all=TRUE) %>%
  select(CityName, State, totalPopulation)

GroupByCity

JointTable <-
  Starbucks %>%
  inner_join(GroupByCity, by=c("City"="CityName", "State"="State"))

Top10TotalPop <-
  JointTable %>%
  arrange(-totalPopulation) %>%
  select(City, Number.of.Starbucks, totalPopulation) %>%
  head(10)

Top10TotalPop

JointTable %>%
  ggplot(aes(x=totalPopulation, y=Number.of.Starbucks)) + geom_point() + ylim(0, 50) + xlim(0, 500000)
```
### Regression Model - Seeing if factors are significant
```{r}
lmStarbucks <- lm(Number.of.Starbucks ~ Median.Age + Median.Household.Income + Percentage.of.white.population
                  + totalPopulation, data=JointTable)
summary(lmStarbucks)
```

### Machine Learning 

```{r}


LeftJointTable <-
  GroupByCity %>%
  left_join(Starbucks, by=c("CityName"="City", "State"="State")) %>%
  filter(State=="California") %>%
  select(Number.of.Starbucks, Median.Age, Percentage.of.white.population, Median.Household.Income, totalPopulation)

LeftJointTable %>%
  arrange(Number.of.Starbucks)

mod1 <- party::ctree(
  Number.of.Starbucks ~ Median.Age + Median.Household.Income + Percentage.of.white.population + totalPopulation,
  data=JointTable)
plot(mod1, type="simple")
```
## Final Comments:
# We can see that there are more Starbucks in cities with younger median age (around 25-40 years old). We don't really see correlations between the number of Starbucks and median household income or percentage of white population. There is a positive correlation between number of Starbucks and the total population in that city; which means that cities with larger populations have more Starbucks. (more people -> more customers -> higher demand -> more profit)

# The decision tree shown has the output Number.of.Starbucks and inputs of Median.Age, Median.Household.Income, Percentage.of.white.population , and totalPopulation. These inputs were the four factors that we were interested in. 
